Fabric-Elasticity Relationships of Tibial Trabecular Bone are Similar in Osteogenesis Imperfecta and Healthy Individuals

### 2021-06-28 09_Article/Simon_2021_Bone_Draft_V01.tex
Correct .gitignore

\begin{abstract}
	Osteogenesis Imperfecta (OI) is an inherited form of bone fragility, also called "brittle bone disease". It is characterised by impaired synthesis of type I collagen, altered trabecular bone architecture and reduced bone mass. High resolution peripheral computed tomography (HR-pQCT) is a powerful method to investigate bone morphology of extremities including the weight-bearing distal tibia. The resulting 3D reconstructions can be used as a basis of micro-finite element (\si{\micro}FE) or homogenised finite element (hFE) models for bone strength estimation. The hFE scheme uses homogenized local bone volume fraction (BV/TV) and anisotropy information (fabric) to compute healthy bone strength within a reasonable computation time using fabric-elasticity relationships. Thus, the aim of this study is to investigate fabric-elasticity relationships in OI trabecular bone compared to healthy controls.

	In this study, the morphology of distal tibiae from 50 adults with OI were compared to 120 healthy controls using second generation HR-pQCT. Six cubic regions of interest (ROIs) were selected per individual in a common anatomical region. A first matching was performed by selecting similar individuals to obtain identical mean and median age and gender distribution in the OI and healthy control group. It allowed to perform a morphometric analysis and compare the outcome with literature. Then, stiffness tensors of ROIs were computed using hFEA and multiple linear regressions were performed on the Zysset-Curnier orthotropic model. The regressions allowed to compare the two groups using 5 parameters. An initial fit was performed on both the OI group and the healthy control group using all the ROIs extracted. Then, data were filtered according to a fixed threshold for a defined coefficient of variation (CV) assessing the ROI heterogeneity and second fits were performed on these filtered data sets. These full and filtered data were in turn compared with previous results from \si{\micro}CT reconstructions obtained in other anatomical locations. Finally, the ROIs of both group were matched according to their BV/TV and fabric anisotropy (DA). Fits were performed again using these matched data to detect statistical differences between the two groups.

@@

Osteogenesis imperfecta (OI), also commonly known as "brittle bone disease", is an inherited form of bone fragility \cite{Tournis2018}. OI prevalence is estimated at about  1/13,500 of births, less severe forms being not accounted in this estimation as they are recognized later in life \cite{Lindahl2015}. Therefore, OI is considered as a rare metabolic bone disorder. In most cases, OI is caused by mutations in genes encoding type I collagen (COL1A1 and COL1A2), leading to brittle and fragile bones \cite{LIM2017}, as well as deformed geometry and size in some cases. OI can be categorized according to disease severity \cite{Mortier2019} into:
\begin{itemize}
	\item Type I: mild
	\item Type II: perinatally lethal
	\item Type III: most severe surviving form
	\item Type IV: intermediate severity
\end{itemize}

>>>>

\begin{abstract}
	Osteogenesis Imperfecta (OI) is an inherited form of bone fragility. It is characterised by impaired synthesis of type I collagen, altered trabecular bone architecture and reduced bone mass. High resolution peripheral computed tomography (HR-pQCT) is a powerful method to investigate bone morphology at peripheral sites including the weight-bearing distal tibia. The resulting 3D reconstructions can be used as a basis of micro-finite element (\si{\micro}FE) or homogenised finite element (hFE) models for bone strength estimation. The hFE scheme uses homogenized local bone volume fraction (BV/TV) and anisotropy information (fabric) to compute healthy bone strength within a reasonable computation time using fabric-elasticity relationships. Thus, the aim of this study is to investigate fabric-elasticity relationships in OI trabecular bone compared to healthy controls.

	In the present study, the morphology of distal tibiae from 50 adults with OI were compared to 120 healthy controls using second generation HR-pQCT. Six cubic regions of interest (ROIs) were selected per individual in a common anatomical region. A first matching between OI and healthy control group was performed by selecting similar individuals to obtain identical mean and median age and gender distribution. It allowed to perform a morphometric analysis and compare the outcome with literature. Then, stiffness tensors of ROIs were computed using hFE and multiple linear regressions were performed with the Zysset-Curnier orthotropic model. The regressions allowed to compare the two groups using five parameters. An initial fit was performed on both the OI group and the healthy control group using all extracted ROIs. Then, data was filtered according to a fixed threshold for a defined coefficient of variation (CV) assessing the ROI heterogeneity and additionnal linear regressions were performed on these filtered data sets. These full and filtered data were in turn compared with previous results from \si{\micro}CT reconstructions obtained in other anatomical locations. Finally, the ROIs of both groups were matched according to their BV/TV and fabric anisotropy (DA). Linear regressions were performed using these matched data to detect statistical differences between the two groups.

@@

Osteogenesis imperfecta (OI), commonly known as brittle bone disease, is an inherited form of bone fragility \cite{Tournis2018}. OI prevalence is estimated at about  1/13,500 of births, less severe forms not being accounted in this estimation as they are recognized later in life \cite{Lindahl2015}. Therefore, OI is considered as a rare metabolic bone disorder. In most cases, OI is caused by mutations in genes encoding type I collagen (COL1A1 and COL1A2), leading to brittle and fragile bones \cite{LIM2017}, as well as deformed geometry and size in some cases. OI can be categorized according to disease severity \cite{Mortier2019}, the most familiar forms are:
\begin{itemize}
	\item Type I: mild
	\item Type II: perinatally lethal
	\item Type III: most severe surviving form
	\item Type IV: intermediate severity
\end{itemize}
